---
title: "Forecast Bi-Annual Electricity Price in Malta"
output: html_notebook
date: Oct 20th, 2019
author: Edward J. Xu (<edxu96@outlook.com>)
---

```{r, include = FALSE}
rm(list = ls())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.asp = 9/16,
  fig.width = 7,
  warning = FALSE
)
```

```{r, warning=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
library(lubridate)
library(forecast)
library(writexl)
library(tseries)
library(magrittr)
```

## 1, Data

```{r, include=FALSE}
ti <-
  readxl::read_excel("~/GitHub/EnergyStorage/e-price-forecast/e-price.xlsx") %>%
  as.matrix() %>%
  t() %>%
  {tibble(bi_annual = as.numeric(rownames(.)), price = .[,1])} %>%
  mutate(observation = rep(TRUE, length(.$bi_annual)))

ti[15:23, 2] <- c(16.47, 16.53, 16.51, 16.66, 16.73, 16.78, 16.64, 16.89, 14.74) / 100
```

Here is the bi-snnual electricity price in Malta (Euro / Kilowatt-hour) with all taxes and levies included, which is downloaded from <https://ec.europa.eu/eurostat/data/database> and <https://www.statista.com/statistics/418104/electricity-prices-for-households-in-malta/>. The data in 2008 and 2009 is missing.

```{r}
head(ti)
```

```{r}
ti %>%
  ggplot() +
    geom_line(mapping = aes(x = bi_annual, y = price)) +
    labs(
      title = "Bi-Annual Electricity Price in Malta",
      subtitle = "with All taxes and levies included from 2003f to 2009s",
      y = "Price (Euro / Kilowatt-hour)",
      x = "Index (Bi-Annual)"
      )
```

## 2, Linear Regression

```{r}
acf(ti$price, na.action = na.pass)
```

```{r}
pacf(ti$price, na.action = na.pass)
```

There is no obvious seasonal variation or fluctuation. So linear regression is used to model the trend. Then, arima model is used to remove the correlation in residuals.

```{r}
lm <- lm(price ~ bi_annual, ti)
```

```{r}
acf(lm$residuals)
```

```{r}
pacf(lm$residuals)
```

## 3, ARIMA Model of Residuals from Linear Regression

```{r}
mod_resi_lm <- auto.arima(lm$residuals)
mod_resi_lm
```

```{r}
acf(mod_resi_lm$residuals)
```

```{r}
pacf(mod_resi_lm$residuals)
```

## 3, Forecast

Forecast is the combination of those from linear regression model and arima model.

```{r}
pred_resi_lm <- predict(mod_resi_lm, n.ahead = 21 * 2)
pred_lm <- predict(
  lm, newdata = data.frame(bi_annual = seq(2019.5, 2019.5 + 20.5, by = 0.5))
  )

ti_fore <- ti %>%
  bind_rows(
    tibble(
      bi_annual = seq(2019.5, 2019.5 + 20.5, by = 0.5),
      price = as.numeric(pred_resi_lm$pred) + pred_lm,
      observation = rep(FALSE, length(pred_lm))
      )
    )
```

```{r}
ti_fore %>%
  ggplot() +
    geom_line(mapping = aes(x = bi_annual, y = price, color = observation)) +
    labs(
      title = "Forecast of Bi-Annual Electricity Price in Malta",
      subtitle = "with all included from 19F to 40F using data from 03f to 09s",
      y = "Price (Euro / Kilo-Watt-Hour)",
      x = "Bi-Annual"
      )
```

## 4, Annual Averaged Price Forecast

```{r}
ti_annual <-
  tibble(
    year = round(seq(2019 - 0.01, 2019 - 0.01 + 20.5, by = 0.5)), 
    price = c(0.1305, (pred_resi_lm$pred + pred_lm)[1:41]), bi = rep(1:2, 21)
    ) %>%
  spread(key = "bi", value = "price") %>%
  mutate(price_ave = (.$`1` + .$`2`) / 2) %>%
  select(year, price_ave)
```

```{r}
ti_annual
```

